Rapid technological progress in white light-emitting diodes \\ and its sources in innovation and technology spillovers

### 2023-08-17 2_SSL_NE/article/article_main.tex
added edits

\hypersetup{
    pdftitle={Main Article},
    pdfauthor={Michael Weinold, Sergey Kolesnikov, Laura Diaz Anadon},
}

@@

M. Weinold$^{1,2,3}$, S. Kolesnikov$^3$, L.D. Anadon$^{3,4}$ \\
\newline
$^1$ CEENRG, Dept. of Land Economy, University of Cambridge, Cambridge,  UK \\
$^2$ Group for Energy Systems Analysis, Department of Mechanical and Process Engineering, ETH Zurich, Zurich, Switzerland \\
$^3$ Technology Assessment Group, Paul Scherrer Institute, Villigen, Switzerland \\
$^4$ Belfer Center for Science and International Affairs, Harvard University, Cambridge MA, USA

@@

\begin{figure}[h!]
 \centering
 \includegraphics[width=\textwidth]{figures/history_efficacy.pdf}
 \caption{\textbf{Historical progress in the luminous efficacy ($K$) of the most widely-used lighting technologies in human history.} Data points indicate best performers by year of market introduction. Luminous efficacy is the measure of how efficiently a light source converts electrical energy into visible light that can be perceived by the human eye, taking into account the wavelength sensitivity of the eye (see Supplementary Note 2 for details). Dashed lines indicate a 3rd-order polynomial fit to the data trend. The physical limit for an ideal light source with a colour rendering index of CRI=90, denoted as $K_{max}^{CRI90}$, is shown as a black horizontal line, as per calculations by Murphy et al. \cite{Murphy2012}. The magnified plot shows the progress in cool white LEDs from 1996 to 2020, with the dashed line indicating a linear rate of efficacy improvement of 10lm/W per year. For comparison, efficacies of best performers in legacy lighting technologies for 2020 are shown as coloured horizontal lines. Note the logarithmic scale of the vertical axis on the main plot and the linear scale on the magnified plot. Abbreviations: HID - High-Intensity Discharge; CFL - Compact Fluorescent Lamp; Hal. - Halogen, Incd. - Incandescent. Source: own synthesis of published data based on a visual approach proposed by Azevedo et al. \cite{azevedo2009transition}. See Supplementary Note 7 for the full list of sources and references.}
 \label{fgr:history_efficacy}
\end{figure}

@@

\begin{figure}[h!]
 \centering
 \includegraphics[width=15.5cm]{figures/waterfall_performance_2003.pdf}
 \includegraphics[width=15.5cm]{figures/waterfall_performance_2010.pdf}
 \includegraphics[width=15.5cm]{figures/waterfall_performance_2016.pdf}
 \includegraphics[width=15.5cm]{figures/waterfall_performance_2020.pdf}
 \captionsetup{font=footnotesize}
 \caption{\textbf{Waterfall diagrams of the loss channels in a generic mid/high-power LED package for 2003, 2010, 2016 and 2020} (top to bottom panels). Losses are normalized to 1 Watt of electric power input (yellow bar on the left). Sub-efficiencies corresponding to each loss channel are listed below each column and described in Supplementary Note 2. Numbers for each loss channel indicate energy losses both in relative terms of input power (in percent) at the point of the channel and absolute values (in Watts). Percentages for loss channels labeled by red, green and blue font indicate losses of corresponding remaining red/green/blue light energy. The following LED architectures and down-conversion phosphor spectra used in calculations in each considered year are shown for reference: 2003 – Flip-chip with YGAG phosphor; 2010 – Flip-chip with 258 phosphor; 2016 – Flip-chip with PFS phosphor; 2020 – Chip-scale package flip-chip with SALON phosphor. Details for each architecture are provided in Figure 3. Details on the phosphors for light down conversion are provided in Table 1 and Supplementary Note 4.  Abbreviations: Scat. = Scattering; Nonrad. = Nonradiative.  Source (Efficiency): own elaboration based on data in Figure SI11. Source (spectral data): Adapted from publications on the respective phosphors: YGAG (2003)\cite{Mueller2002}, 258 (2010)\cite{MuellerMach2005}, PFS (2016)\cite{Murphy2015}, SALON (2019)\cite{Hoerder2019}.}
 \label{fgr:waterfall}
\end{figure}

@@

\begin{table}[h!]
    \footnotesize
    \centering
    \caption{\textbf{LED innovations related to improvements in consumer experience metrics.}}
    \begin{NiceTabularX}{\textwidth}{|l|l|l|X|X|l|}
    \hline
        \textbf{Year} & \textbf{Desig.} & \textbf{Chemical formula} & \textbf{Description} & \textbf{Significance} & \textbf{Ref.} \\ \hline
        1996 & YAG & $Y_3 Al_5 O_{12}:Ce$ & Yttrium aluminium garnet (YAG) phosphor activated with cerium & First LED phosphor, enabled white LEDs & SI 3.2.2.1 \\ \hline
        1996 & YGAG & $(Y_{1-x} Gd_x)_3 Al_5 O_{12}:Ce$ & Gadolinium-doped YAG phosphor & First red-shifted phosphor, enabled warm white LEDs & SI 3.2.2.1 \\ \hline
        2002 & 258 & $(Ba,Sr)_2 Si_5 N_8:Eu^{2+}$ & Europium-doped nitridosilicate phosphor & First red LED phosphor & SI 3.2.2.2 \\ \hline
        2003 & QD & N/A & Quantum dot-based phosphor & First use of QD for LED light down conversion & SI 3.2.2.4 \\ \hline
        2005 & PFS & $K_2 SiF_6: Mn^{4+}$ & Manganese-activated potassium fluorosilicate (PFS) phosphor & First ultra-narrow-band red LED phosphor & SI 3.2.2.3 \\ \hline
        2013 & SLA & $Sr[Li Al]_3 N_4 ]:Eu^{2+}$ & Europium-doped cuboidal nitridolithoaluminate phosphor & Improved narrow-band red phosphor & SI 3.2.2.2 \\ \hline
        2016 & SALON & $Sr[Li_2 Al_2 O_2 N_2]:Eu^{2+}$ & Europium-doped oxonitride phosphor & High-performance ultra-narrow-band red phosphor & SI 3.2.2.2 \\ \hline
    \end{NiceTabularX}
    \caption{\textit{Year} column represents the earliest reported application of innovation in white LEDs. These differ from the years shown in \cref{fgr:consumer_experience} which correspond to the earliest publication of spectral data for LED products that relied on those innovations. Detailed descriptions of the history of each innovation and relevant spectral data are provided in Supplementary Note 4. Desig. = Designation, Ref. - Reference.}
    \label{tab:phosphors}
\end{table}

@@

    \caption{Note: Disc. - Year of initial discovery, identified  by the earliest literature source describing the discovery of original idea or invention outside the LED domain. S/O - Year of spillover to LED; Comm. - Year of commercial application, identified as the year of the first recorded application of that idea or invention in a commercial LED product. Ref. - References. LED innovations are ordered by the year in which a technology spillover into LED occurred, provided in the S/O column. Origin column represents knowledge domains in which spillovers initially emerged, where (S) denotes a scientific discipline and (T) is an area of technology. Ref. column lists literature sources for the represented discoveries, innovations and spillovers. Area of Improvement column represents the impact of spillovers on different aspects of white LED technology, e.g., improvements in particular sub-efficiencies.}

@@

@@

Our LED manufacturing cost model shows a 95.5\% decrease in the cost of producing low-to-mid power GaN-based white classic-chip LEDs from 1.11\$ to 0.05\$ (in 2020 USD) between 2003 and 2020. In contrast with LED performance improvements, where progress was driven mostly by innovations in LED technology, the dramatic decline in LED manufacturing cost has been a product of higher yields across manufacturing steps and economies of scale resulting from increases in the sapphire wafer size. Both were enabled by advances in manufacturing equipment (e.g., for epitaxy and chip dicing) and incremental manufacturing process improvements from learning-by-doing. Surprisingly, LED innovations and associated technology spillovers are not found to have a substantial impact on LED manufacturing cost reductions over time. [PLACEHOLDER TO DISCUSS THE REASONS FOR THAT].

@@

\section{Conclusions}

Our study has analysed the sources of the dramatic cost reductions and performance improvements in white light-emitting diodes since their introduction to the market in 1996 in a systematic and granular way. We find that the total LED device efficiency increase from 5.8\% to 38.8\% between 2003 and 2020, as well as improvements in consumer experience metrics, have been predominantly driven by LED technology innovations that affected all physical energy loss channels and corresponding device sub-efficiencies. We also find that among those innovations, at least nine were driven by knowledge spillovers originating in areas of science and technology beyond LEDs or solid-state lighting. These spillover-driven innovations were responsible for 8.5\% of the total efficiency improvements and nearly 100\% of the improvements in consumer experience metrics.

Our manufacturing cost model shows a 95.5\% decrease in the cost of producing white classic-chip LEDs (from 1.11\$(2020) to 0.05\$(2020)) between 2003 and 2020, driven mostly by increases in the wafer size and yields across different manufacturing steps over time. In contrast with performance improvements, these cost reductions were thus a result of economies of scale and learning by doing,  which were in turn likely facilitated by policies creating and growing market demand for LED lighting, such as incandescent light bulb bans and subsidies for LEDs.

Our analysis of the sources, mechanisms and enablers of the identified technology spillovers which were significant drivers of improvements in LED efficiency and consumer experience metrics, highlights the critical role played by a deep understanding of the physical, chemical and optical phenomena underlying the operation of LEDs, as well as materials science and technology and nanotechnology involved in the production of LEDs, for past and future advances in LED and solid-state lighting technology. Specifically, deep physical understanding of LED device efficiency loss channels enabled important innovations in LEDs that increased several sub-efficiencies in LEDs and will continue to do so, as expected by eminent experts in the field \cite{Weisbuch2020}. This suggests that additional research in these areas and a more deliberate search for relevant external knowledge may accelerate expected future advances in LED technology. These knowledge spillovers can be enabled or even accelerated by knowledge exchange events and long-term partnerships between academia and industry, dedicated mission-driven public R\&D funding, and a freedom of search in academia. This further reinforces arguments made against the dichotomy of basic research versus applied research \cite{narayanamurti2016cycles,narayanamurti2021genesis} and the calls for open, inclusive and flexible research cultures \cite{Stephan2021}.

There are various important avenues of future research that are opened up by our analysis. First, future work could expand the cost model by collecting and including data for a broader set of chip architectures and analyzing the impact of individual innovations and spillovers on the costs as opposed to performance. Second, a deeper dive on the role of learning-by-doing is needed both in the cost and performance analysis. Third, building on the work on LED sub-efficiencies and physical limits, future efforts could focus on identifying priority areas for further efficiency improvements in LEDs and SSL in general. Finally, by comparing the drivers of innovation, technology spillovers, cost reductions and performance improvements at a granular level across different clean energy technologies, we can identify patterns or differences that would help us formulate recommendations for industry and policymakers aimed at accelerating further clean energy innovation for climate change mitigation.

>>>>

\hypersetup{
    pdftitle={Main Article},
    pdfauthor={Michael Philipp Weinold, Sergey Kolesnikov, Laura Diaz Anadon},
}

@@

M. P. Weinold$^{1,2,3}$, S. Kolesnikov$^1$, L.D. Anadon$^{1,4}$ \\
\newline
$^1$ Centre for Environment, Energy and Natural Resource Governance, Department of Land Economy, University of Cambridge, Cambridge, UK \\
$^2$ Group for Energy Systems Analysis, Department of Mechanical and Process Engineering, ETH Zurich, Zurich, Switzerland \\
$^3$ Technology Assessment Group, Paul Scherrer Institute, Villigen, Switzerland \\
$^4$ Belfer Center for Science and International Affairs, Harvard University, Cambridge MA, USA

@@

\begin{figure}[h!]
 \centering
 \includegraphics[width=\textwidth]{figures/history_efficacy.pdf}
 \caption{\textbf{Historical progress in the luminous efficacy ($\eta$) of the most widely-used lighting technologies in human history.} Data points indicate best performers by year of market introduction. Luminous efficacy is the measure of how efficiently a light source converts electrical energy into visible light that can be perceived by the human eye, taking into account the wavelength sensitivity of the eye (see Supplementary Note 2 for details). Dashed lines indicate a 3rd-order polynomial fit to the data trend. The physical limit for an ideal light source with a colour rendering index of CRI=90, denoted as $K_{max}^{CRI90}$, is shown as a black horizontal line, as per calculations by Murphy et al. \cite{Murphy2012}. The magnified plot shows the progress in cool white LEDs from 1996 to 2020, with the dashed line indicating a linear rate of efficacy improvement of 10lm/W per year. For comparison, efficacies of best performers in legacy lighting technologies for 2020 are shown as coloured horizontal lines. Note the logarithmic scale of the vertical axis on the main plot and the linear scale on the magnified plot. Abbreviations: HID - High-Intensity Discharge; CFL - Compact Fluorescent Lamp; Hal. - Halogen, Incd. - Incandescent. Source: own synthesis of published data based on a visual approach proposed by Azevedo et al. \cite{azevedo2009transition}. See Supplementary Note 7 for the full list of sources and references.}
 \label{fgr:history_efficacy}
\end{figure}

@@

\begin{figure}[h!]
 \centering
 \includegraphics[width=15.5cm]{figures/waterfall_performance_2003.pdf}
 \includegraphics[width=15.5cm]{figures/waterfall_performance_2010.pdf}
 \includegraphics[width=15.5cm]{figures/waterfall_performance_2016.pdf}
 \includegraphics[width=15.5cm]{figures/waterfall_performance_2020.pdf}
 \captionsetup{font=footnotesize}
 \caption{\textbf{Waterfall diagrams of the loss channels in a generic mid/high-power LED package for 2003, 2010, 2016 and 2020} (top to bottom panels). Losses are normalized to 1 Watt of electric power input (yellow bar on the left). Sub-efficiencies corresponding to each loss channel are listed below each column and described in Supplementary Note 2.  Numbers for each loss channel indicate energy losses both in relative terms of input power (in percent) at the point of the channel and absolute values (in Watts). Percentages for loss channels labeled by red, green and blue font indicate losses of corresponding remaining red/green/blue light energy. The following LED architectures and down-conversion phosphor spectra used in calculations in each considered year are shown for reference: 2003 – Flip-chip with YGAG phosphor; 2010 – Flip-chip with 258 phosphor; 2016 – Flip-chip with PFS phosphor; 2020 – Chip-scale package flip-chip with SALON phosphor. Details for each architecture are provided in \cref{fgr:chip_architecture_overview}.  Details on the phosphors for light down conversion are provided in \cref{tab:phosphors} and Supplementary Note 4. Abbreviations: Scat. = Scattering; Nonrad. = Nonradiative. Source: own elaboration based on data on sub-efficiencies presented in Supplementary Figure 9 and spectral data in Supplementary Figure 10,  adapted from publications on the respective phosphors: YGAG (2003)\cite{Mueller2002}, 258 (2010)\cite{MuellerMach2005}, PFS (2016)\cite{Murphy2015}, SALON (2019)\cite{Hoerder2019}.}
 \label{fgr:waterfall}
\end{figure}

@@

\begin{table}[h!]
    \footnotesize
    \centering
    \caption{\textbf{LED innovations related to improvements in consumer experience metrics.}}
    \begin{NiceTabularX}{\textwidth}{|l|l|l|X|X|l|}
    \hline
        \textbf{Year} & \textbf{Desig.} & \textbf{Chemical formula} & \textbf{Description} & \textbf{Significance} & \textbf{Ref.} \\ \hline
        1996 & YAG & $Y_3 Al_5 O_{12}:Ce$ & Yttrium aluminium garnet (YAG) phosphor activated with cerium & First LED phosphor, enabled white LEDs & SI 3.2.2.1 \\ \hline
        1996 & YGAG & $(Y_{1-x} Gd_x)_3 Al_5 O_{12}:Ce$ & Gadolinium-doped YAG phosphor & First red-shifted phosphor, enabled warm white LEDs & SI 3.2.2.1 \\ \hline
        2002 & 258 & $(Ba,Sr)_2 Si_5 N_8:Eu^{2+}$ & Europium-doped nitridosilicate phosphor & First red LED phosphor & SI 3.2.2.2 \\ \hline
        2003 & QD & N/A & Quantum dot-based phosphor & First use of QD for LED light down conversion & SI 3.2.2.4 \\ \hline
        2005 & PFS & $K_2 SiF_6: Mn^{4+}$ & Manganese-activated potassium fluorosilicate (PFS) phosphor & First ultra-narrow-band red LED phosphor & SI 3.2.2.3 \\ \hline
        2013 & SLA & $Sr[Li Al]_3 N_4 ]:Eu^{2+}$ & Europium-doped cuboidal nitridolithoaluminate phosphor & Improved narrow-band red phosphor & SI 3.2.2.2 \\ \hline
        2016 & SALON & $Sr[Li_2 Al_2 O_2 N_2]:Eu^{2+}$ & Europium-doped oxonitride phosphor & High-performance ultra-narrow-band red phosphor & SI 3.2.2.2 \\ \hline
    \end{NiceTabularX}
    \caption*{Note: \textit{Year} column represents the earliest reported application of innovation in white LEDs. These differ from the years shown in \cref{fgr:consumer_experience} which correspond to the earliest publication of spectral data for LED products that relied on those innovations. Detailed descriptions of the history of each innovation and relevant spectral data are provided in Supplementary Note 4. Desig. = Designation, Ref. - Reference.}
    \label{tab:phosphors}
\end{table}

@@

    \caption*{Note: Disc. - Year of initial discovery, identified  by the earliest literature source describing the discovery of original idea or invention outside the LED domain. S/O - Year of spillover to LED; Comm. - Year of commercial application, identified as the year of the first recorded application of that idea or invention in a commercial LED product. Ref. - References. LED innovations are ordered by the year in which a technology spillover into LED occurred, provided in the S/O column. Origin column represents knowledge domains in which spillovers initially emerged, where (S) denotes a scientific discipline and (T) is an area of technology. Ref. column lists literature sources for the represented discoveries, innovations and spillovers. Area of Improvement column represents the impact of spillovers on different aspects of white LED technology, e.g., improvements in particular sub-efficiencies.}

@@

In this study, we use a multi-method approach to synthesize evidence from multiple sources and, for the first time, reconstruct a comprehensive picture of the rapid technological progress in white LED technology throughout its history since the introduction to the market in 1996 and across an ensemble of metrics related to device performance, cost, and consumer experience. Improvements in these metrics are traced to sources in innovation, technology spillovers, and economies of scale in the manufacturing process.

@@

Our LED manufacturing cost model shows a 95.5\% decrease in the cost of producing low-to-mid power GaN-based white classic-chip LEDs from 1.11\$ to 0.05\$ (in 2020 USD) between 2003 and 2020. In contrast with LED performance improvements, where progress was driven mostly by innovations in LED technology, the dramatic decline in LED manufacturing cost has been a product of higher yields across manufacturing steps and economies of scale resulting from increases in the sapphire wafer size. Surprisingly, LED innovations and associated technology spillovers are not found to have a substantial impact on LED manufacturing cost reductions over time. There are two reasons for this:

First, the \textit{functional unit} of our study is the LED chip itself, not the luminaire containing multiple chips. As efficiency and overall LED brightness has increased, a smaller number of chips could be used in luminaires to obtain the same level of overall flux. However, since we are investigating improvements at the level of single chips, we do not capture cost effects downstream of chip manufacturing. Nonetheless, we were previously able to show that overall brightness is no longer a target metric in chip development \cite{weinold2021compound}.

Second, we find that an increase in wafer size and higher overall yield together are responsible for a 98\% reduction of manufacturing cost per LED die between between 2003 and 2020. Both were enabled by advances in manufacturing equipment (e.g., for epitaxy and chip dicing) and incremental manufacturing process improvements from learning-by-doing, which was confirmed by multiple interviews with experts from manufacturing. A plausible explanation for this finding is that the main focus of R&D activities in white LED technology in the considered period was likely on LED performance improvements rather than cost reductions.

@@

### 2023-08-17 2_SSL_NE/article/article_supplementary.tex
updated custom names
some more edits
added edits

% nature energy custom code
\makeatletter
\renewcommand{\@seccntformat}[1]{Supplementary Note \csname the#1\endcsname\quad}
\makeatother

@@

% custom figure and table numbering
\renewcommand{\thefigure}{Supplementary Figure \arabic{figure}}
\renewcommand{\thetable}{SI\arabic{table}}

% custom equation numbering
% https://tex.stackexchange.com/a/184641/
\newcounter{defcounter}
\setcounter{defcounter}{0}
\newenvironment{myequation}{%
\addtocounter{equation}{-1}
\refstepcounter{defcounter}
\renewcommand\theequation{SI\thedefcounter}
\begin{equation}}
{\end{equation}}

@@

\hypersetup{
    pdftitle={Supplement},
    pdfauthor={Michael Weinold, Sergey Kolesnikov, Laura Diaz Anadon},
}

@@

\setlength{\fboxsep}{10pt}
\fbox{
    \parbox{\textwidth}{
        \textit{\textsc{Supplementary Information}} for: \\
        \newline 
        M. Weinold$^{1,2,3}$, S. Kolesnikov$^3$, L.D. Anadon$^{3,4}$ \\
        "Rapid technological progress in white light-emitting diodes and its sources in innovation and technology spillovers" (2023) \\
        \newline
        $^1$ Technology Assessment Group, Paul Scherrer Institute, Switzerland \\
        $^2$ Group for Energy Systems Analysis, Department of Mechanical and Process Engineering, ETH Zurich, Zurich, Switzerland \\
        $^3$ CEENRG, Dept. of Land Economy, University of Cambridge, UK \\
        $^4$ Belfer Center for Science and International Affairs, Harvard University, Cambridge MA, USA
    }
}

\tableofcontents

\section{Previous Literature on Technological Progress in LEDs}
\label{sec:prev_lit}

@@

Due to these limitations, any recommendations on future improvements in LED technology based on these studies are fragmented at best. Furthermore, the effect of individual innovations and technology spillovers, which has been investigated in solar PV \cite{kavlak2018evaluating,kolesnikov2020novel,nemet2019solar}, and to some extent in lithium-ion batteries \cite{Stephan_2021}, has not yet been studied in the context of lighting.

@@

LED light-extraction efficiency (LEE) describes the  losses due to absorption in the material after electron-hole recombination and the associated emission of a photon. It is defined as
%
\begin{equation}
    \eta_{LE} = \frac{P_{opt}}{P_{int}} = \frac{\text{\# of photons out/s}}{\text{\# of photons created/s}}
\end{equation}
%
where $P_{opt}$ is the optical power emitted from the device $P_{int}$ is the optical power emitted from the active region into other parts of the device (internal) \cite{schubert2018light}\cite{tsao2010solid}.

@@

Light conversion efficiency (LCE) describes the sum of all losses in the process of converting the blue light that is the basis for all phosphor-converted white LEDs. Specifically, energy is lost due to three distinct effects. First, the inherent wavelength-conversion loss or \textit{Stokes shift} as the difference between the original blue wavelength and the down-converted red, yellow or green wavelength. Second, non-radiative processes in the phosphor material. Third, scattering and absorption processes in the phosphor material. These three effects are listed individually in \cref{fgr:waterfall}. The overall light conversion efficiency is defined as
%
\begin{equation}
    \eta_{C} = \frac{E_{\textcolor{blue}{B}}}{\sum_{i=\textcolor{red}{R},\textcolor{orange}{O},\textcolor{yellow}{Y},\textcolor{teal}{G}} E_i}
\end{equation}
%
where $E_i$ denotes the total energy of light at color $i$ corresponding to the down-converted photon wavelength \cite{schubert2018light}\cite{tsao2010solid}. Since every individual phosphor component in the device has its own associated conversion losses, the denominator sums over all phosphor components, i.e. \textcolor{red}{R}ed, \textcolor{orange}{O}range, \textcolor{yellow}{Y}ellow, \textcolor{teal}{G}reen.

@@

Spectral efficiency (SE) describes losses in the conversion process due to the wavelength-dependent efficiency of the human eyes. Photons converted into the infrared or ultraviolet are lost to illumination purposes. It is defined as
%
\begin{equation}
    \eta_{S} = \frac{K}{K_{max}(CRI,CCT)}
\end{equation}
%
where $K$ is the luminous efficacy of radiation of the light source, which can be computed from the device spectrum and the luminosity function, which describes the sensitivity of the human eye. $K_{max}$ is the maximum luminous efficacy of radiation of a perfect light source with the same color rendering performance and correlated color temperature as the light source in question\cite{schubert2018light}\cite{tsao2010solid}.

@@

The luminous efficacy \textit{of radiation} $K$ describes the match of a light-emitting diode package spectrum to the human visual system. Efficacy in lighting is dependent on the luminosity function, which describes the wavelength-dependent sensitivity of the human eye. A light source emitting very \textit{efficiently} in the infrared yet emitting no visible light has a very low \textit{efficacy}. The luminous efficacy of radiation $K$ is mathematically defined as the normalized, integrated product of the spectral radiant flux of a light source with the wavelength-dependent human sensitivity to light \cite{cie-term-effrad}
%
\begin{myequation}
\label{eqn:ler}
    K [\text{lm/W}_{opt}]= \frac{\int_0^\infty K( \lambda ) \phi \text{d} \lambda}{\int_0^\infty \phi \text{d} \lambda}
\end{myequation}
%
where
%
\begin{align*}
    K &\dots \text{luminosity function} \\
    \phi &\dots \text{spectral radiant flux} \\
    \lambda &\dots \text{wavelength}
\end{align*}
%
This metric can be computed from spectral data alone and does not require additional spectral normalization. It enables straightforward comparison between the performance of different downconversion phosphors, as shown in the top panel of \cref{fig:phosphor_spectrum}. Light sources emitting in the far red or blue part of the spectrum have lower efficacy of radiation.

Care must be taken not to confuse this metric with luminous efficacy of source described in \cref{subsec:les}.

@@

The luminous efficacy \textit{of a light source} $\eta$ is defined as the ratio between the emitted luminous flux and the consumed electrical power $P_{el}$ \cite{cie-term-effsrc}. It is often used to describe general progress in lighting technologies.
%
\begin{myequation}
    \eta [\text{lm/W}_{el}]= \frac{\phi}{P_{el}}
\end{myequation}
%
This metric is also frequently cited in device datasheets, scientific literature and textbooks when describing the performance of light-emitting diodes. However, as noted in \cref{subsec:ler}, care must be taken not to confuse this efficacy metric with the \textit{luminous efficacy of radiation}, which depends only on the spectral characteristics of a light source. As the luminous efficacy of a light source $\eta$ captures the overall device efficacy, it depends on a large number of other device properties and parameters. This makes attribution of changes in this metric to individual changes in device design or manufacturing difficult. For this reason, we do not use this metric only to show historical progress across different lighting technologies in Figure 1 in the main text of the article.

@@

The colour rendering index (CRI) of a light source describes its ability to render the colours of an object faithfully when compared to illumination under a reference light source, such as standard daylight \cite{khan2015led}. The way it is calculated is defined by the International Illumination Commission (CIE) \cite{cie_cri_1995}. CRI has certain limitations when applied to solid-state light sources \cite{david2013cri}. \cite{Houser2013}, CRI has remained the de facto industry standard for describing colour fidelity of light sources \cite{DOE2016LED}. High colour rendering performance of lighting is a requirement in workplace environments, retail stores, clinical operating environments and art exhibitions \cite{khanh2017color}. It should be noted that some niche applications prioritize high colour saturation over high CRI, for instance, in food display or fabric retail applications \cite{david2013cri}. However, these niche applications remain outside our focus on general illumination. Due to a broad availability and importance of CRI data for consumers of various LED lighting sources, we adopt CRI as the key metric to track progress in consumer experience in LED lighting, despite its limitations.

>>>>

% cross-referencing
\input{2_SSL_NE/article/auxiliary/preamble/1_preamble_functionality_crossreference}

% custom section designations
\makeatletter
\renewcommand{\@seccntformat}[1]{Supplementary Note \csname the#1\endcsname: } % https://tex.stackexchange.com/a/693618
\makeatother
% custom equation designations
\renewcommand{\figurename}{Supplementary Figure}
\crefname{figure}{Supplementary Figure}{Supplementary Figures}
\renewcommand{\tablename}{Supplementary Table}
\crefname{table}{Supplementary Table}{Supplementary Tables}

@@

@@

\hypersetup{
    pdftitle={Supplement},
    pdfauthor={Michael Philipp Weinold, Sergey Kolesnikov, Laura Diaz Anadon},
}

@@

\setlength{\fboxsep}{10pt}
\fbox{
    \parbox{\textwidth}{
        \textit{\textsc{Supplementary Information}} for: \\
        \newline 
        M. P. Weinold$^{1,2,3}$, S. Kolesnikov$^1$, L.D. Anadon$^{1,4}$ \\
        \newline
        $^1$ Centre for Environment, Energy and Natural Resource Governance, Department of Land Economy, University of Cambridge, Cambridge, UK \\
        $^2$ Group for Energy Systems Analysis, Department of Mechanical and Process Engineering, ETH Zurich, Zurich, Switzerland \\
        $^3$ Technology Assessment Group, Paul Scherrer Institute, Villigen, Switzerland \\
        $^4$ Belfer Center for Science and International Affairs, Harvard University, Cambridge MA, USA
    }
}

\clearpage
\tableofcontents

\clearpage
\section{Previous Literature on Techological Progress in LEDs}
\label{sec:prev_lit}

@@

Due to these limitations, any recommendations on future improvements in LED technology based on these studies are fragmented at best. Furthermore, the effect of individual innovations and technology spillovers, which has been investigated in solar PV \cite{kavlak2018evaluating,kolesnikov2020novel,nemet2019solar}, and to some extent in lithium-ion batteries \cite{Stephan2021}, has not yet been studied in the context of lighting.

@@

LED light extraction efficiency (LEE) describes the  losses due to absorption in the material after electron-hole recombination and the associated emission of a photon. It is defined as
%
\begin{equation}
    \eta_{LE} = \frac{P_{opt}}{P_{int}} = \frac{\text{\# of photons out/s}}{\text{\# of photons created/s}}
\end{equation}
%
where $P_{opt}$ is the optical power emitted from the device and  $P_{int}$ is the optical power emitted from the active region into other parts of the device (internal) \cite{schubert2018light}\cite{tsao2010solid}.

@@

Light conversion efficiency (LCE) describes the sum of all losses in the process of converting the blue light that is the basis for all phosphor-converted white LEDs. Specifically, energy is lost due to three distinct effects. The first effect is the inherent wavelength-conversion loss or \textit{Stokes shift} as the difference between the original blue wavelength and the down-converted red, yellow or green wavelength. The second effect are non-radiative processes in the phosphor material. The third , scattering and absorption processes in the phosphor material. These three effects are listed individually in Figure 4. The overall light conversion efficiency is defined as
%
\begin{equation}
    \eta_{C} = \frac{E_{\textcolor{blue}{B}}}{\sum_{i=\textcolor{red}{R},\textcolor{orange}{O},\textcolor{yellow}{Y},\textcolor{teal}{G}} E_i}
\end{equation}
%
where $E_{\textcolor{blue}{B}}$ is the total energy of blue light before down-conversion and $E_i$ denotes the total energy of light at color $i$ corresponding to the down-converted photon wavelength \cite{schubert2018light}\cite{tsao2010solid}. Since every individual phosphor component in the device has its own associated conversion losses, the denominator sums over all phosphor components, i.e. \textcolor{red}{R}ed, \textcolor{orange}{O}range, \textcolor{yellow}{Y}ellow, \textcolor{teal}{G}reen.

@@

Spectral efficiency (SE) describes losses in the light down conversion process due to the wavelength-dependent efficiency of the human eyes. Photons converted into the infrared or ultraviolet light are lost to illumination purposes. It is defined as
%
\begin{equation}
    \eta_{S} = \frac{K}{K_{max}(CRI,CCT)}
\end{equation}
%
where $K$ is the luminous efficacy of radiation of the light source, defined in \cref{eqn:ler}, which can be computed from the device spectrum and the luminosity function, which describes the sensitivity of the human eye. $K_{max}$ is the maximum luminous efficacy of radiation of a perfect light source with the same color rendering index performance and correlated color temperature as the light source in question\cite{schubert2018light}\cite{tsao2010solid}.

@@

The luminous efficacy \textit{of radiation} $K$ is a supplementary metric used in our calculations of SE, rather than directly in our analysis of the improvements in LED performance. Care must be taken not to confuse this metric with the luminous efficacy of a light source, described in \cref{subsec:les}.

The luminous efficacy \textit{of radiation} $K$ describes the match of the spectrum of a light-emitting diode to the human visual system. Efficacy in lighting is dependent on the luminosity function, which describes the wavelength-dependent sensitivity of the human eye. A light source emitting very \textit{efficiently} in the infrared yet emitting no visible light has a very low \textit{efficacy}. The luminous efficacy of radiation $K$ is mathematically defined as the normalized, integrated product of the spectral radiant flux of a light source $\phi$ with the wavelength-dependent human sensitivity to light \cite{cie-term-effrad}
%
\begin{equation}
\label{eqn:ler}
    K [\text{lm/W}_{opt}]= \frac{\int_0^\infty K( \lambda ) \phi \text{d} \lambda}{\int_0^\infty \phi \text{d} \lambda}
\end{equation}
%
where
%
\begin{align*}
    K &\dots \text{luminosity function} \\
    \phi &\dots \text{spectral radiant flux} \\
    \lambda &\dots \text{wavelength}
\end{align*}
%
This metric can be computed from spectral data alone and does not require additional spectral normalization. It enables straightforward comparison between the performance of different downconversion phosphors, as shown in the top panel of \cref{fig:phosphor_spectrum}. Light sources emitting in the far red or blue part of the spectrum have lower efficacy of radiation.

@@

The luminous efficacy \textit{of a light source} $\eta$ is defined as the ratio between the emitted luminous flux $\phi$ and the consumed electrical power $P_{el}$ \cite{cie-term-effsrc}.
%
\begin{equation}
    \eta [\text{lm/W}_{el}]= \frac{\phi}{P_{el}}
\end{equation}
%
This metric is also frequently cited in device datasheets, scientific literature and textbooks when describing the performance of light-emitting diodes. However, as noted in \cref{subsec:ler}, care must be taken not to confuse this efficacy metric with the \textit{luminous efficacy of radiation}, which depends only on the spectral characteristics of a light source. As the luminous efficacy of a light source $\eta$ captures the overall device efficacy, it depends on a large number of other device properties and parameters. This makes attribution of changes in this metric to individual changes in device design or manufacturing difficult. For this reason, we do not use this metric only to show historical progress across different lighting technologies in Figure 1 in the main text of the article.

@@

The colour rendering index (CRI) of a light source describes its ability to render the colours of an object faithfully when compared to illumination under a reference light source, such as standard daylight \cite{khan2015led}. The way it is calculated is defined by the International Illumination Commission (CIE) \cite{cie_cri_1995}. CRI has certain limitations when applied to solid-state light sources \cite{david2013cri}. However, despite repeated attempts at constructing more elaborate colour rendering metrics \cite{Houser2013}, CRI has remained the de facto industry standard for describing colour fidelity of light sources \cite{DoE2016LED}. High colour rendering performance of lighting is a requirement in workplace environments, retail stores, clinical operating environments and art exhibitions \cite{khanh2017color}. It should be noted that some niche applications prioritize high colour saturation over high CRI, for instance, in food display or fabric retail applications \cite{david2013cri}. However, these niche applications remain outside our focus on general illumination. Due to a broad availability and importance of CRI data for consumers of various LED lighting sources, we adopt CRI as the key metric to track progress in consumer experience in LED lighting, despite its limitations.
